---
title: "Naive DD"
subtitle: "Scoring, M2 Data Science for Social Sciences"
author: "Théo Druilhe, Pierre Larose, Nathan Pizzetta, Sigurd Saue"
format: 
  html:
    theme: 
       light: cerulean
    code-fold: false
    warning: false
    message: false
    toc: true
    toc-depth: 3
    embed-resources: true
    self-contained: true
    code-tools:
      source: true
editor: 
    markdown: 
       wrap: sentence
---

```{r, echo=FALSE}
library(dplyr)
library(readr)
library(DT)
library(RPostgres)
library(lubridate)
```




## 1. Calculate the market value of assets (V)
The market value of assets is estimated as the sum of the market value of equity and total debt

```{r}
df = read_csv("data.csv")
data <- df %>%
  mutate(
    market_value_equity = market_to_book * book_value, # Market value of equity
    total_assets = book_value,                         # Approximate total assets using book value
    V = market_value_equity + (leverage * book_value)  # Market value of assets = equity + debt
  )
```



## 2. Calculate total debt (D)
Total debt is derived using the leverage ratio and book value of assets

```{r}
data <- data %>%
  mutate(
    D = leverage * book_value # Total debt
  )
```


## 3. Calculate the Naive Distance to Default (Naive DD)
Naive DD = 1 - (Debt / Market value of assets)

```{r}
data <- data %>%
  mutate(
    naive_DD = 1 - (D / V)
  )
```


## 4. Select relevant columns to display or save the results

```{r}
naive_DD_results <- data %>%
  select(fyear, gvkey, market_value_equity, D, V, naive_DD)
```



## 5. Merge with Y column and filter
```{r}
naive_DD_results <- naive_DD_results %>%
  left_join(data %>% select(fyear, gvkey, Y), by = c("fyear", "gvkey"))

naive_DD_filtered <- naive_DD_results %>%
  filter(Y == 1)
```

### Results dataset :
```{r, echo = FALSE}
datatable(naive_DD_results, 
          options = list(pageLength = 10,
                         autoWidth = TRUE,
                         searchable = TRUE, 
                         filter = "top"))
```
### Results dataset (With only Y = 1)
```{r, echo = FALSE}
datatable(naive_DD_filtered, 
          options = list(pageLength = 10,
                         autoWidth = TRUE,
                         searchable = TRUE, 
                         filter = "top"))
```