---
title: "practical1_task2"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(survival)
library(dplyr)
```

## Data Manipulation

Import the data

```{r}
setwd("/Users/theodruilhe/Documents/M2_D3S/scoring_project")

# Import X
file_path <- file.path("data", "our_data", "X.rds")
if (file.exists(file_path)) {
  X <- readRDS(file_path)
} else {
  print("File not found. Please check the file path.")
}

# import target_Y
file_path <- file.path("data", "our_data", "target_Y.rds")
if (file.exists(file_path)) {
  y <- readRDS(file_path)
} else {
  print("File not found. Please check the file path.")
}
```

Create an new ID for the join

```{r}
X <- X %>%
  mutate(ID = paste(gvkey, fyear, sep = "_"))
y <- y %>%
  mutate(ID = paste(gvkey, fyear, sep = "_"))
```

Join the two data set

```{r}
# Perform the inner join
data <- y %>%
  inner_join(X, by = c("ID" = "ID"))
```

Delete the useless columns

```{r}
data <- data %>%
  select(-fyear.y, -gvkey.y)
```

Rename gvkey and fyear

```{r}
colnames(data)[colnames(data) == "gvkey.x"] <- "gvkey"
colnames(data)[colnames(data) == "fyear.x"] <- "fyear"
```

Transform Y as a binary variable

```{r}
data <- data %>%
  mutate(Y = ifelse(Y == 1, 1, 0))
```

## Descriptive Statistics

```{r}
# get info about the data
str(data)
```

We apply the same technique as Shumway to deal with extreme values. Specifically: 1. Values above the 99th percentile of each variable are capped (set) to the 99th percentile value. 2. Values below the 1st percentile of each variable are floored (set) to the 1st percentile value.

```{r}
# Function to truncate values at 1st and 99th percentiles
truncate_outliers <- function(column) {
  p1 <- quantile(column, 0.01, na.rm = TRUE)  # 1st percentile
  p99 <- quantile(column, 0.99, na.rm = TRUE) # 99th percentile
  column <- ifelse(column < p1, p1, column)   # Floor at 1st percentile
  column <- ifelse(column > p99, p99, column) # Cap at 99th percentile
  return(column)
}

# Apply truncation to all numeric columns except Y
data <- data %>%
  mutate(across(where(is.numeric) & !all_of("Y"), truncate_outliers))
```

Missing values processing

```{r}
# Check for missing values and give the proportion of missing values by column
missing_values <- data %>%
  summarise(across(everything(), ~ sum(is.na(.)) / n())) %>%
  gather() %>%
  arrange(desc(value))

# Print missing values
print(missing_values)
```

```{r}
# install.packages("naniar")
library(naniar)
gg_miss_var(data)  # Plot missing values by variable
```

Due to the big number of missing values in these two variables we will delete them

```{r}
summary(data$interest_coverage)
```

```{r}
summary(data$inventory_turnover)
```

```{r}
summary(data$book_value)
```

```{r}
summary(data$roic)
```

Here we affect the value of the median for all Na

```{r}
summary(data$ebitda_margin)
```

```{r}
# affect the median value to all NA's
data$roic[is.na(data$roic)] <- median(data$roic, na.rm = TRUE)

data$ebitda_margin[is.na(data$ebitda_margin)] <- median(data$ebitda_margin, na.rm = TRUE)

data$free_cash_flow_to_sales[is.na(data$free_cash_flow_to_sales)] <- median(data$free_cash_flow_to_sales, na.rm = TRUE)

data$leverage[is.na(data$leverage)] <- median(data$leverage, na.rm = TRUE)

data$operating_cash_flow_to_debt[is.na(data$operating_cash_flow_to_debt)] <- median(data$operating_cash_flow_to_debt, na.rm = TRUE)

data$debt_to_equity[is.na(data$debt_to_equity)] <- median(data$debt_to_equity, na.rm = TRUE)

data$debt_ratio[is.na(data$debt_ratio)] <- median(data$debt_ratio, na.rm = TRUE)

data$PE_ratio[is.na(data$PE_ratio)] <- median(data$PE_ratio, na.rm = TRUE)

data$ROE[is.na(data$ROE)] <- median(data$ROE, na.rm = TRUE)

data$ROA[is.na(data$ROA)] <- median(data$ROA, na.rm = TRUE)

data$asset_turnover[is.na(data$asset_turnover)] <- median(data$asset_turnover, na.rm = TRUE)

data$market_to_book[is.na(data$market_to_book)] <- median(data$market_to_book, na.rm = TRUE)

data$net_profit_margin[is.na(data$net_profit_margin)] <- median(data$net_profit_margin, na.rm = TRUE)
```

```{r}
data_cleaned <- data %>%
  select(-book_value, -inventory_turnover, -interest_coverage, -ebitda)
```

```{r}
gg_miss_var(data_cleaned)  # Plot missing values by variable
```

## Survival analysis (Hazard Model)

```{r}
#install.packages("survival")
library(survival)
```

Create time and status variables

```{r}
# Prepare survival data
df <- data_cleaned %>%
  group_by(gvkey) %>% # Group by firm identifier
  mutate(time = fyear - min(fyear),  # Calculate time-to-event
         status = Y) %>%            # Status (1 = bankruptcy, 0 = no event)
  ungroup()
```

Descriptive Statistics

```{r}
# make descriptive statistics of data
str(df)
```

Fit the Cox Proportional Hazards Model (all the variables except ones define below with values goes to infinity)

```{r}
# Fit the Cox Proportional Hazards Model
cox_model <- coxph(Surv(time, status) ~ 
                     ROA + ROE + net_profit_margin + asset_turnover + 
                     debt_to_equity + debt_ratio + PE_ratio + 
                     market_to_book + operating_cash_flow_to_debt + 
                     free_cash_flow_to_sales + ebitda_margin + roic + leverage, 
                   data = df)

# View model summary
summary(cox_model)
```

### Model evaluation

The C-index measures the model’s ability to rank survival times correctly.

```{r}
# Concordance index
cox_summary <- summary(cox_model)
c_index <- cox_summary$concordance[1]  # First element of concordance contains the C-index
cat("Concordance Index (C-Index):", c_index, "\n")
```

risk scores (linear predictors) to use them for stratification or classification

```{r}
# Compute risk scores
risk_scores <- predict(cox_model, type = "risk")

# Add risk scores to the dataset
df <- df %>%
  mutate(risk_score = risk_scores)
```

If we dichotomize the risk (e.g., high vs. low risk based on the median risk score), you can compute confusion matrices.

```{r}
# Dichotomize risk scores: high risk (1) vs. low risk (0) based on the median
threshold <- median(risk_scores)
df <- df %>%
  mutate(predicted_status = ifelse(risk_score > threshold, 1, 0))

# Create confusion matrix
library(caret)
confusion <- confusionMatrix(as.factor(df$predicted_status), as.factor(df$status))
print(confusion)
```

## Logistic Model

```{r}
library(tidyverse)
library(broom)
```

```{r}
logistic_model <- glm(
  Y ~ ROA + ROE + net_profit_margin + asset_turnover + 
       debt_to_equity + debt_ratio + PE_ratio + 
       market_to_book + operating_cash_flow_to_debt + 
       free_cash_flow_to_sales + ebitda_margin + roic + leverage, 
  family = binomial(link = "logit"),
  data = data_cleaned
)

# Step 3: Summarize the results
summary(logistic_model)
```

```{r}
# Load necessary libraries
library(tidyverse)
library(caret)     # For train-test split
library(pROC)      # For evaluation

# Step 1: Prepare the dataset
# Ensure the dataset is named 'data_cleaned' and has no missing values or issues.
data <- data_cleaned

# Step 2: Train-test split
set.seed(123)  # Set seed for reproducibility
train_index <- createDataPartition(data$Y, p = 0.7, list = FALSE)  # 70% training data
train_data <- data[train_index, ]
test_data <- data[-train_index, ]

# Step 3: Train the Logistic Regression Model
logistic_model <- glm(
  Y ~ ROA + ROE + net_profit_margin + asset_turnover + 
       debt_to_equity + debt_ratio + PE_ratio + 
       market_to_book + operating_cash_flow_to_debt + 
       free_cash_flow_to_sales + ebitda_margin + roic + leverage, 
  data = train_data
)

# Step 4: Summarize the Model
summary(logistic_model)

# Step 5: Evaluate the Model on Test Data
# Predict probabilities for the test set
test_data <- test_data %>%
  mutate(predicted_prob = predict(logistic_model, newdata = ., type = "response"))

# Create predictions based on a cutoff (e.g., 0.5)
test_data <- test_data %>%
  mutate(predicted_class = ifelse(predicted_prob > 0.5, 1, 0))

# Step 6: Model Evaluation Metrics
# Confusion Matrix
confusion_matrix <- table(test_data$predicted_class, test_data$Y)
print("Confusion Matrix:")
print(confusion_matrix)

# Accuracy
accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)
print(paste("Accuracy:", round(accuracy, 4)))

# ROC Curve and AUC
roc_curve <- roc(test_data$Y, test_data$predicted_prob)
print(paste("AUC:", round(auc(roc_curve), 4)))
plot(roc_curve, col = "blue", main = "ROC Curve")

# Precision, Recall, F1 Score (optional)
precision <- confusion_matrix[2, 2] / sum(confusion_matrix[2, ])
recall <- confusion_matrix[2, 2] / sum(confusion_matrix[, 2])
f1_score <- 2 * (precision * recall) / (precision + recall)
print(paste("Precision:", round(precision, 4)))
print(paste("Recall:", round(recall, 4)))
print(paste("F1 Score:", round(f1_score, 4)))
```

## Task 3:

Classic **K-fold Cross-Validation** randomly splits the data into k folds, which is appropriate for independent and identically distributed (i.i.d.) data. However, time series data like yours (structured as firm-year observations) often exhibits **temporal dependency**: outcomes in one period can depend on past periods. Random splits would violate the temporal ordering and could lead to **data leakage**, where future information influences training.

Instead, **Time Series Cross-Validation (Walk Forward Validation)** respects the temporal structure by ensuring that each training set includes only observations up to the prediction period. This method simulates a real-world scenario where only past data is available for forecasting.

```{r}
# Step 1: Prepare data
data <- data_cleaned

# Step 2: Sort data by firm (gvkey) and year (fyear) to respect temporal structure
data <- data %>% arrange(gvkey, fyear)

# Step 3: Define Time Series Cross-Validation (Walk Forward Scheme)
# Create custom indices for training and testing
time_series_cv <- function(data, n_splits) {
  indices <- list()
  n <- nrow(data)
  split_size <- floor(n / (n_splits + 1))  # Calculate split size
  
  for (i in 1:n_splits) {
    train_end <- split_size * i  # End index for training
    test_start <- train_end + 1
    test_end <- test_start + split_size - 1
    
    if (test_end > n) break  # Ensure test indices stay within range
    
    train_indices <- seq(1, train_end)
    test_indices <- seq(test_start, test_end)
    
    indices[[i]] <- list(train = train_indices, test = test_indices)
  }
  indices
}

# Create 5 splits for time series CV
n_splits <- 5
cv_indices <- time_series_cv(data, n_splits)
```

```{r}

# Step 4: Perform Walk Forward Validation
results <- list()

for (i in seq_along(cv_indices)) {
  # Get train and test data
  train_data <- data[cv_indices[[i]]$train, ]
  test_data <- data[cv_indices[[i]]$test, ]
  
  # Fit logistic regression model on training data
  logistic_model <- glm(
    Y ~ ROA + ROE + net_profit_margin + asset_turnover + 
       debt_to_equity + debt_ratio + PE_ratio + 
       market_to_book + operating_cash_flow_to_debt + 
       free_cash_flow_to_sales + ebitda_margin + roic + leverage, 
    family = binomial(link = "logit"),
    data = train_data
  )
  
  # Predict probabilities on test data
  test_data <- test_data %>%
    mutate(predicted_prob = predict(logistic_model, newdata = ., type = "response"),
           predicted_class = ifelse(predicted_prob > 0.5, 1, 0))
  
  # Evaluate model performance
  confusion_matrix <- table(test_data$predicted_class, test_data$Y)
  accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)
  roc_curve <- roc(test_data$Y, test_data$predicted_prob)
  auc <- auc(roc_curve)
  
  # Store results
  results[[i]] <- list(
    fold = i,
    accuracy = accuracy,
    auc = auc,
    confusion_matrix = confusion_matrix,
    roc_curve = roc_curve
  )
}
```

```{r}
# Step 5: Summarize Results
accuracy_list <- sapply(results, function(x) x$accuracy)
auc_list <- sapply(results, function(x) x$auc)

cat("Walk Forward Validation Results:\n")
cat("Average Accuracy:", round(mean(accuracy_list), 4), "\n")
cat("Average AUC:", round(mean(auc_list), 4), "\n")
```

```{r}
# Plot the last ROC curve as an example
plot(results[[n_splits]]$roc_curve, col = "blue", main = "ROC Curve for Last Fold")
```
